---
title: "4.4 - Soil Depth Analyses"
author: "Vicki M. Zhang"
date: "April 17, 2025"
output: html_document
---
# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warnings = F)
```

```{r}
# packages
library(here)
library(tidyverse)
library(ggmap)
library(ggsn)
library(sp)
library(ggstar)
library(lme4)
library(lmerTest)
library(car)
library(MuMIn)
library(emmeans)
library(multcomp)
library(ggfortify)
library(patchwork)
library(cowplot)
```


## Data


```{r directory, include = F}
depthperf_wide <- read_csv(here("4.4_Soil_Depth/data/depthperf_wide.csv"))
depthseed_wide <- read_csv(here("4.4_Soil_Depth/data/depthseed_wide.csv"))
```

### Wrangling Performance

```{r}
df_depth <- depthperf_wide %>% 
  mutate_at(c("site", "habitat", "plot", "species"), as.factor) %>% 
  mutate(treatment = factor(treatment, levels = c("0", "5", "20"))) %>% 
  mutate_at(c("survival23", "dead23", "survival24", "dead24"), as.integer)
```

```{r survival}
df_survival <- df_depth %>% 
  pivot_longer(cols = c(survival23, survival24),
               names_to = "year_survival",
               names_prefix = "survival",
               values_to = "survival") %>% 
  pivot_longer(cols = c(dead23, dead24),
               names_to = "year_dead",
               names_prefix = "dead",
               values_to = "dead") %>% 
  dplyr::select(c(habitat, site, treatment, species, year_survival, survival, year_dead, dead)) %>% 
  # remove duplicate years
  filter(year_survival == year_dead) %>% 
  mutate_at(c("year_survival", "year_dead"),
            as.numeric) %>% 
  mutate("year_survivalf" = as.factor(year_survival))

df_lvul_survival <- df_survival %>% 
  filter(species == "Lvul")

df_pmaj_survival <- df_survival %>% 
  filter(species == "Pmaj")

df_toff_survival <- df_survival %>% 
  filter(species == "Toff")
```

```{r performance long}
df_performance <- df_depth %>%
  mutate(across(c(biomass_a24, biomass_b24, biomass24), ~. * 1000)) %>%  # change biomass to mg
  dplyr::select(site, habitat, treatment, species,
                height23, rhizome23, leaflength23, taproot23,
                height24, ramets24, leaflength24, numleaves24, numflower24, flower24,
                biomass_a24, biomass_b24, biomass24) %>% 
  pivot_longer(cols = -c(site, habitat, treatment, species), # initial measurements pivoted 
               names_to = c("performance", "year"),
               names_pattern = "([a-z]+)([0-9]+)",
               values_to = "values") %>% 
  mutate_at("year", as.numeric)

df_performance_init <- df_depth %>%
  mutate(across(c(biomass_a24, biomass_b24, biomass24), ~. * 1000)) %>%  # change biomass to mg
  dplyr::select(site, habitat, treatment, species,
                height23, rhizome23, leaflength23, taproot23,
                height24, ramets24, leaflength24, numleaves24, numflower24, flower24,
                biomass_a24, biomass_b24, biomass24) %>% 
  pivot_longer(cols = -c(site, habitat, treatment, species,
                         height23, rhizome23, leaflength23, taproot23), # initial measurements kept in separate column
               names_to = c("performance", "year"),
               names_pattern = "([a-z]+)([0-9]+)",
               values_to = "values") %>% 
  mutate_at("year", as.numeric)

df_lvul_performance <- df_performance_init %>% 
  filter(species == "Lvul")

df_pmaj_performance <- df_performance_init %>% 
  filter(species == "Pmaj")

df_toff_performance <- df_performance_init %>% 
  filter(species == "Toff")
```


### Wrangling Seed

```{r}
depthseed_wide <- depthseed_wide %>% 
  mutate_at(c("site", "habitat", "plot", "treatment"),
            as.factor)
```


```{r seed bank long}
df_seed <- depthseed_wide %>% 
  pivot_longer(cols = c("Pmaj":"Gtet"),
               names_to = "species",
               values_to = "seed") %>% 
  mutate_at("species", as.factor)
```


# Map

```{r churchill on canada map}
churchill <- data.frame(longitude = c(-94.1650), latitude = c(58.7684))

map_insetcanada <- map_data("world", region = "canada") %>% 
  ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "white", colour = "black") +
  geom_star(data = churchill,
            aes(x = longitude , y = latitude),
            size = 3,
            fill = "black") +
  theme(line = element_blank(),
        axis.title = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        plot.background = element_blank())

# ggsave(here::here("4.4_Soil_Depth/plots/map_insetcanada.png"), map_insetcanada,
#        width = 8, height = 6)
```

```{r transplant sites coordinates}
# 16 unique sites
sites_depth <- df_depth %>%
  distinct(site, habitat, lat, lon) # select only one instance of each site

# add common garden field experiment
site_field <- data.frame(habitat = "CG",
                         site = 9,
                         lat = 58.73416667,
                         lon = -93.79138889)

df_sites <- rbind(sites_depth, site_field)
```

```{r sites}
map_churchill <- get_googlemap(center = "Malcolm Ramsay Lake, MB, Canada",
                         zoom = 12,
                         style = c(feature = "all",
                                   element = "labels", visibility = "off"),
                         maptype = "hybrid",
                         color = "bw", force = TRUE)

map_mainchurchill <- ggmap(map_churchill) +
  geom_point(data = df_sites,
             aes(x = lon, y = lat, fill = habitat, shape = habitat),
            size = 3) +
  scale_fill_brewer(palette = "Accent",
                    labels = c("Boreal", "Tundra", "Common Garden")) +
  scale_shape_manual(values = c(21, 24, 22),
                     labels = c("Boreal", "Tundra", "Common Garden")) +
  scalebar(location = "bottomright",
           x.min = -93.7, x.max = -93.75,
           y.min = 58.67, y.max = 58.7,
           dist = 2, dist_unit = "km",
           transform = TRUE, model = 'WGS84', 
           st.dist = 0.05, st.size = 3, st.color = "white",
           border.size = 0.5) +
  labs(y = "Longitude", x = "Latitude") +
  theme(panel.border = element_rect(colour = "black", fill=NA))

# ggsave(here("4.4_Soil_Depth/plots/map_mainchurchill.png"), map_mainchurchill, width = 8, height = 6)
```

```{r inset and main}
map_final <- map_mainchurchill +
  inset(ggplotGrob(map_insetcanada),
        xmin = -93.78, xmax = -93.68,
        ymin = 58.735, ymax = 58.775)

# ggsave(here("4.4_Soil_Depth/plots/map_final.png"), map_final, width = 8, height = 7)
```

```{r soil collection sites}
site_collections <- data.frame(habitat = c("Granary Ponds", "Port Roads", "Cape Merry Roads", "Dump Entrance"),
                         site = c(1, 2, 3, 4),
                         lat = c(58.7775, 58.781667, 58.7788889, 58.7575),
                         lon = c(-94.191389, -94.195556, -94.18333333333334, -93.9816666667))

map_churchill <- get_googlemap(center = "Churchill Airport, Manitoba",
                         zoom = 11,
                         style = c(feature = "all",
                                   element = "labels", visibility = "off"),
                         maptype = "hybrid",
                         color = "bw", force = TRUE)

map_maincollection <- ggmap(map_churchill) +
  geom_point(data = site_collections,
             aes(x = lon, y = lat, shape = habitat),
             size = 2, fill = "red") +
  scale_shape_manual(name = "Soil Collection Sites", values = c(21, 22, 24, 23)) +
  scalebar(location = "bottomright",
           x.min = -94.0, x.max = -93.9,
           y.min = 58.68, y.max = 58.69,
           dist = 2, dist_unit = "km",
           transform = TRUE, model = 'WGS84',
           st.dist = 0.5, st.size = 3, height = 0.1, st.color = "white",
           border.size = 0.2) +
  labs(y = "Longitude", x = "Latitude") +
  theme(panel.border = element_rect(colour = "black", fill=NA))

# ggsave(here("4.4_Soil_Depth/plots/map_mainchurchill.png"), map_maincollection, width = 8, height = 6)
```


# Survival

```{r survival per species summary}
survival_sum <- df_depth %>% 
  pivot_longer(cols = c(survival23, survival24),
               names_to = "status",
               values_to = "survival") %>% 
  separate_wider_position(cols = "status",
                          widths = c(status = 8, year = 2)) %>% 
  group_by(year, species, habitat, treatment) %>%  
  dplyr::summarise(alive = sum(survival == 1, na.rm = TRUE),
                   dead = sum(survival == 0, na.rm = TRUE),
                   propsurvival = alive/(alive + dead),
                   sd = sd(survival),
                   se = sd/sqrt(8))
```

```{r total survival per species}
fig_sppsurvival <- survival_sum %>%
  filter(year == "24") %>% 
  ggplot(aes(x = treatment, y = propsurvival, fill = habitat, shape = habitat,
             group = interaction(habitat, species))) +
  facet_wrap(~ species,
             labeller = as_labeller(c("Lvul" = "L. vulgaris",
                                      "Pmaj" = "P. major",
                                      "Toff" = "T. officinale")),
             scales = "free_y",
             strip.position = "left") + 
  geom_errorbar(aes(x = treatment, ymin = propsurvival - se, ymax = propsurvival + se),
                position = position_dodge(width = 0.5),
                colour = "black") +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Proportion of surviving transplants",
                     limits = c(0, 1), expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_brewer(palette = "Accent",
                      name = "Habitat",
                      labels = c("Boreal", "Tundra")) +
  scale_shape_manual(name = "Habitat",
                     labels = c("Boreal", "Tundra"),
                     values = c(21, 22)) + 
  theme_vz() +
  # theme_darkclassic() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(face = "italic"),
        # legend.position = "none"
        )


# ggsave(here::here("4.4_Soil_Depth/plots/fig_sppsurvival.png"), fig_sppsurvival,
#        width = 12, height = 7)
```


# Biomass

```{r}
fig_sppbiomass <- df_performance %>% 
  group_by(habitat, treatment, species) %>%
  filter(!is.na(values) & performance == "biomass") %>% 
  dplyr::summarise(mean = mean(values),
                   n = n(),
                   sd = sd(values),
                   se = sd/sqrt(n)) %>% 
  ggplot(aes(x = treatment, y = mean,
             fill = habitat)) +
  facet_wrap(~ species,
             labeller = as_labeller(c("Lvul" = "L. vulgaris",
                                      "Pmaj" = "P. major",
                                      "Toff" = "T. officinale")),
             scales = "free_y",
             strip.position = "left") + 
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single")) +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(width = 0.9),
                colour = "black") +
  scale_x_discrete(name = "Soil depth (cm)") +
  scale_y_continuous(name = "Mean biomass (mg)",
                     limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  theme_vz() +
  # theme_darkclassic() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(face = "italic"),
        # legend.position = "none"
        )

# ggsave(here::here("4.4_Soil_Depth/plots/fig_sppbiomass.png"), fig_sppbiomass,
#        width = 12, height = 7)
```

# Flower

```{r}
flower_sum <- df_performance %>% 
  group_by(habitat, treatment, species) %>% 
  filter(!is.na(values) & performance == "flower") %>% 
  dplyr::summarise(total = sum(values),
                   mean = mean(values),
                   n = n(),
                   sd = sd(values),
                   se = sd/sqrt(n))
```

```{r}
flower_sum %>% 
  ggplot(aes(x = treatment, y = mean, colour = habitat,
             shape = species)) +
  facet_wrap(~ species) +
  geom_point(position = position_dodge2(width = 0.9)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge2(0.9)) +
  scale_colour_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  theme_vz()

# ggsave(here::here("4.4_Soil_Depth/plots/fig_sppflower.png"), fig_sppflower,
#        width = 12, height = 7)
```


# Soil Seed Bank

## Summary

```{r}
seed_sum <- df_seed %>% 
  group_by(species, habitat, treatment) %>% 
  filter(species != "Salb") %>% # no germinants
  droplevels() %>% 
  summarize(total = sum(seed, na.rm = TRUE),
            mean = mean(seed, na.rm = TRUE),
            sd = sd(seed, na.rm = TRUE),
            se = sd/sqrt(total))
```

```{r}
# histogram including NA
df_seed %>% 
  group_by(seed) %>%
  count(seed) %>%
  ggplot(aes(x = as.factor(seed), y = n)) + 
  geom_col()

# histogram shows right-skew
df_seed %>% 
  ggplot(aes(x = seed)) + 
  geom_histogram(bins = 10) + 
  facet_wrap(~ species)
```


## GLM

```{r}
seed_fixed <- glm(seed ~ treatment + habitat + species,
                  family = poisson(link = "log"),
                  data = df_seed)

seed_fixed1 <- glm(seed ~ treatment * habitat + species,
                   family = "poisson",
                   data = df_seed)

seed_mixed <- glmer(seed ~ treatment + habitat + species + (1|site),
                    family = "poisson",
                    data = df_seed)

seed_mixed1 <- glmer(seed ~ treatment * habitat + species + (1|site),
                     family = "poisson",
                     data = df_seed)

summary(seed_fixed)
Anova(seed_fixed, p.adjust.methods = TRUE, type = 3, test.statistic = "Wald")

# model comparison
AICc(seed_fixed, seed_fixed1, seed_mixed, seed_mixed1)

# post-hoc
seed_pairs <- emmeans(seed_fixed, ~ treatment + habitat | species, type = "response")
summary(seed_pairs)
pairs(seed_pairs, adjust = "tukey")
```

## Figure

```{r point by species}
fig_seedbank <- ggplot(seed_sum, aes(x = treatment,
                     y = total,
                     shape = habitat,
                     colour = species,
                     group = interaction(habitat, species))) +
  geom_jitter(size = 2,
              position = position_dodge(width = 0.5)) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Number of emerging seedings") +
  scale_colour_brewer(palette = "Dark2",
                      name = "Species",
                      labels = c(expression(italic("C. bursa-pastoris")),
                                 expression(italic("G. tetrahit")),
                                 expression(italic("P. major")),
                                 expression(italic("T. arvense")),
                                 expression(italic("T. officinale"))), 
                     # values = c(1:5)
                     ) +
  scale_shape_manual(name = "Habitat",
                     labels = c("Boreal", "Tundra"),
                     values = c(15, 16)) + 
  theme_vz() +
  # theme_darkclassic() +
  guides(shape = guide_legend(override.aes = list(colour = "black"))) +
  theme(legend.text.align = 0,
        # legend.position = "none"
        )
```

```{r}
fig_seedbank <- ggplot(seed_sum) +
  geom_col(aes(x = treatment, y = total,
               fill = habitat),
           colour = "black",
           position = position_dodge2(preserve = "single")) +
  facet_wrap(~species,
             labeller = as_labeller(c("Cbp" = "C. bursa-pastoris",
                                      "Gtet" = "G. tetrahit",
                                      "Pmaj" = "P. major",
                                      "Tarv" = "T. arvense",
                                      "Toff" = "T. officinale")),
             scales = "fixed",
             axes = "all_y",
             strip.position = "left",
             nrow = 1) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Total seed bank germination",
                     limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_brewer(name = "Habitat", palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(face = "italic"),
        # legend.position = "none"
        )

# ggsave(here::here("4.4_Soil_Depth/plots/fig_seedbank.png"), fig_seedbank,
#        width = 12, height = 7)
```




## PCA

```{r for PCA}
df_seedpca <- depthseed_wide %>% 
  unite(treat, c("site", "habitat", "treatment"), sep = "_") %>% 
  mutate_at(c("treat"), as.factor) %>% 
  dplyr::select(-c(2:4)) %>% 
  column_to_rownames(var = "treat") %>% 
  replace(is.na(.), 0)
```


<!-- ```{r} -->
<!-- seed_pca <- prcomp(x = df_seedpca, center = TRUE, scale = TRUE) -->

<!-- summary(seed_pca) -->
<!-- seed_pca$rotation -->
<!-- seed_pca$x -->
<!-- plot(seed_pca) -->

<!-- autoplot(seed_pca, loadings = TRUE, loadings.label = TRUE, -->
<!--          data = depthseed_wide, colour = "treatment") -->
<!-- ``` -->

```{r}
library(PerformanceAnalytics)
df_seedpca %>% 
  chart.Correlation()
```



# Individual Species Performance

## *L. vulgaris

### Survival

```{r total survival}
# figure
fig_lvultotalsurvival <- survival_sum %>% 
  group_by(habitat, treatment) %>% 
  filter(species == "Lvul") %>% 
  
  # plot
  ggplot(aes(x = treatment,
             y = propsurvival,
             fill = habitat)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste("n =", alive)),
            position = position_dodge(0.9),
            vjust = -0.5) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Total number of surviving transplants") +
  scale_fill_brewer(name = "Soil Origin",
                    palette = "Accent",
                    labels = c("Dump", "Tundra")) +
  theme_vz()

# ggsave(here::here("4.4_Soil_Depth/plots/fig_lvultotalsurvival.png"), fig_lvultotalsurvival,
#        width = 12, height = 7)
```

#### GLM

```{r}
# glm
survival_fixed <- glm(survival ~ treatment + habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Lvul", ],
                      subset = year_survivalf != "23")

survival_fixed1 <- glm(survival ~ treatment * habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Lvul", ],
                      subset = year_survivalf != "23")

summary(survival_fixed)
Anova(survival_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")

survival_mixed <- glmer(survival ~ treatment + habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Lvul", ],
                        subset = year_survivalf != "23")

survival_mixed1 <- glmer(survival ~ treatment * habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Lvul", ],
                        subset = year_survivalf != "23")

survival_mixed2 <- glmer(survival ~ habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Lvul", ],
                        subset = year_survivalf != "23")

survival_mixed3 <- glmer(survival ~ treatment + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Lvul", ],
                        subset = year_survivalf != "23")


# model comparison
AICc(survival_fixed, survival_fixed1, survival_mixed, survival_mixed1, survival_mixed2, survival_mixed3)
```

```{r}
survival_mixed <- glm(survival ~ treatment + habitat,
                        family = "binomial",
                        data = df_survival[df_survival$species == "Lvul", ],
                        subset = year_survivalf != "23")

summary(survival_mixed)
Anova(survival_mixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")


# post-hoc
surv_pairs <- emmeans(survival_fixed, ~ treatment + habitat,
                      type = "response")
summary(surv_pairs)
cld(surv_pairs, Letters = letters)

pairs(surv_pairs, adjust = "tukey", reverse = "TRUE")
```

### Performance

```{r normality}
# normality of height
shapiro.test(df_lvul_performance[df_lvul_performance$performance == "height", ]$values)
hist(df_lvul_performance[df_lvul_performance$performance == "height", ]$values)

# normality of ramets
shapiro.test(df_lvul_performance[df_lvul_performance$performance == "ramets", ]$values)
hist(df_lvul_performance[df_lvul_performance$performance == "ramets", ]$values)

# normality of biomass
shapiro.test(df_lvul_performance[df_lvul_performance$performance == "biomass", ]$values)
hist(df_lvul_performance[df_lvul_performance$performance == "biomass", ]$values)
```

```{r transformation}
df_lvul_performance$values_transformed <- log(df_lvul_performance$values)
```

```{r Lvul performance summary}
lvul_perf_sum <- df_lvul_performance %>% 
  group_by(habitat, treatment, species, performance) %>% 
  filter(!is.na(values) &
           (performance == "height" | performance == "ramets" | performance == "biomass")) %>% 
  summarize(mean = mean(values),
            sd = sd(values),
            n = n(),
            se = sd/sqrt(n)) %>%
  mutate(mean = replace_na(mean, 0),
         se = replace_na(se, 0))
```

#### LMM

Need to use transformed values for height and ramets.

```{r height models}
# lm
lvul_fixed1 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment * habitat + height23 + rhizome23)

lvul_fixed2 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment + habitat + height23 + rhizome23)

lvul_fixed3 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment * habitat + height23)

lvul_fixed4 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment + habitat + height23)

lvul_fixed5 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment * habitat + rhizome23)

lvul_fixed <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment + habitat + rhizome23)

# lmer
lvul_mixed1 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment * habitat + height23 + rhizome23 + (1|site),
                   REML = F)

lvul_mixed2 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment + habitat + height23 + rhizome23 + (1|site),
                   REML = F)

lvul_mixed3 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment * habitat + height23 + (1|site),
                   REML = F)

lvul_mixed4 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment + habitat + rhizome23 + (1|site),
                   REML = F)

lvul_mixed5 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment * habitat + rhizome23 + (1|site),
                   REML = F)

lvul_mixed <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                   values_transformed ~ treatment + habitat + height23 + (1|site),
                   REML = F)

# comparison
AICc(lvul_fixed, lvul_fixed1, lvul_fixed2, lvul_fixed3, lvul_fixed4, lvul_fixed5,
     lvul_mixed, lvul_mixed1, lvul_mixed2, lvul_mixed3, lvul_mixed4, lvul_mixed5)
```

```{r final model}
lvul_fixed <- lm(data = df_lvul_performance[df_lvul_performance$performance == "height", ],
                 values_transformed ~ treatment + habitat + rhizome23)

# normality
qqnorm(residuals(lvul_fixed))
qqline(residuals(lvul_fixed))

shapiro.test(residuals(lvul_fixed))

# coefficients
summary(lvul_fixed)
Anova(lvul_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(lvul_fixed)

# post-hoc
lvul_emmeans <- emmeans(lvul_fixed, ~ treatment + habitat,
                        type = "response")

summary(lvul_emmeans)
lvul_emmeans1 <- cld(lvul_emmeans, Letters = letters) # significance letters

pairs(lvul_emmeans, adjust = "tukey")
contrast(lvul_emmeans, simple = "treatment")
```

```{r ramets models}
# lm
lvul_fixed3 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment * habitat + height23 + rhizome23)

lvul_fixed1 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment + habitat + height23 + rhizome23)

lvul_fixed2 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment * habitat + height23)

lvul_fixed <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment + habitat + height23)

lvul_fixed4 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment * habitat + rhizome23)

lvul_fixed5 <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment + habitat + rhizome23)

# lmer
lvul_mixed1 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment * habitat + height23 + rhizome23 + (1|site),
                   REML = F)

lvul_mixed2 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment + habitat + height23 + rhizome23 + (1|site),
                   REML = F)

lvul_mixed3 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment * habitat + height23 + (1|site),
                   REML = F)

lvul_mixed <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment + habitat + height23 + (1|site),
                   REML = F)

lvul_mixed4 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment * habitat + rhizome23 + (1|site),
                   REML = F)

lvul_mixed5 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                   values_transformed ~ treatment + habitat + rhizome23 + (1|site),
                   REML = F)

# comparison
AICc(lvul_fixed, lvul_fixed1, lvul_fixed2, lvul_fixed3, lvul_fixed4, lvul_fixed5,
     lvul_mixed, lvul_mixed1, lvul_mixed2, lvul_mixed3, lvul_mixed4, lvul_mixed5)
```


```{r final model}
lvul_fixed <- lm(data = df_lvul_performance[df_lvul_performance$performance == "ramets", ],
                 values_transformed ~ treatment + habitat + height23)

# normality
qqnorm(residuals(lvul_fixed))
qqline(residuals(lvul_fixed))

shapiro.test(residuals(lvul_fixed))

# coefficients
summary(lvul_fixed)
Anova(lvul_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(lvul_fixed)

# post-hoc
lvul_emmeans <- emmeans(lvul_fixed, ~ treatment + habitat,
                        type = "response")

summary(lvul_emmeans)
lvul_emmeans2 <- cld(lvul_emmeans, Letters = letters) # significance letters

pairs(lvul_emmeans, adjust = "tukey")
contrast(lvul_emmeans, simple = "treatment")
```

#### Figure

```{r Lvul performance plot}
# add performance labels
lvul_emmeans1$performance <- "height"
lvul_emmeans2$performance <- "ramets"

# bind emmeans of performance measurements
lvul_emmeans <- rbind(lvul_emmeans1, lvul_emmeans2)

# bind to summary
emm_df <- full_join(lvul_perf_sum[lvul_perf_sum$performance != "biomass", ],
                    lvul_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)


fig_lvulperf <- lvul_perf_sum %>% 
  filter(performance == "height" | performance == "ramets") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single"),
           colour = "black") +
  facet_wrap(~ performance,
             labeller = as_labeller(c("height" = "Mean height (cm)",
                                      "ramets" = "Mean number of ramets")),
             scales = "free_y",
             strip.position = "left") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL,
                     limits = c(0, NA), expand = expansion(mult = c(0, 0.3))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
  position = position_dodge2(width = 0.9)) +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  theme_vz() +
  # theme_darkclassic() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none"
        )

# ggsave(here::here("4.4_Soil_Depth/plots/fig_lvulperf.png"), fig_lvulperf,
#        width = 12, height = 7)
```


### Flower

```{r flower models}
# lm
lvul_fixed1 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment * habitat + height23 + rhizome23)

lvul_fixed2 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment + habitat + height23 + rhizome23)

lvul_fixed3 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment * habitat + height23)

lvul_fixed4 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment + habitat + height23)

lvul_fixed5 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment * habitat + rhizome23)

lvul_fixed <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment + habitat + rhizome23)

# lmer
lvul_mixed1 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment * habitat + height23 + rhizome23 + (1|site),
                  REML = F)

lvul_mixed2 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                    values ~ treatment + habitat + height23 + rhizome23 + (1|site),
                    REML = F)

lvul_mixed3 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                   values ~ treatment * habitat + height23 + (1|site),
                   REML = F)

lvul_mixed4 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                    values ~ treatment + habitat + height23 + (1|site),
                    REML = F)

lvul_mixed5 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                    values ~ treatment * habitat + rhizome23 + (1|site),
                    REML = F)

lvul_mixed <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                    values ~ treatment + habitat + rhizome23 + (1|site),
                    REML = F)

# comparison
AICc(lvul_fixed, lvul_fixed1, lvul_fixed2, lvul_fixed3, lvul_fixed4, lvul_fixed5,
     lvul_mixed, lvul_mixed1, lvul_mixed2, lvul_mixed3, lvul_mixed4, lvul_mixed5)
```

```{r final model}
lvul_fixed <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "flower", ],
                  values ~ treatment + habitat + rhizome23)

summary(lvul_fixed)
Anova(lvul_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(lvul_mixed)

# post-hoc
lvul_emmeans <- emmeans(lvul_mixed, ~ treatment + habitat,
                        type = "response")
summary(lvul_emmeans)
lvul_emmeans3 <- cld(lvul_emmeans, Letters = letters) # significance letters

lvul_pairs <- pairs(lvul_emmeans, adjust = "tukey")
```


### Biomass

log-transformed

```{r biomass models}
# lm
lvul_fixed1 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + height23 + rhizome23)

lvul_fixed2 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + height23 + rhizome23)

lvul_fixed3 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + height23)

lvul_fixed <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + height23)

lvul_fixed5 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + rhizome23)

lvul_fixed4 <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + rhizome23)

# lmer
lvul_mixed1 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + height23 + rhizome23 + (1|site),
                  REML = F)

lvul_mixed2 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                    values_transformed ~ treatment + habitat + height23 + rhizome23 + (1|site),
                    REML = F)

lvul_mixed3 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                   values_transformed ~ treatment * habitat + height23 + (1|site),
                   REML = F)

lvul_mixed4 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                    values_transformed ~ treatment + habitat + height23 + (1|site),
                    REML = F)

lvul_mixed5 <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                    values_transformed ~ treatment * habitat + rhizome23 + (1|site),
                    REML = F)

lvul_mixed <- lmer(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                    values_transformed ~ treatment + habitat + rhizome23 + (1|site),
                    REML = F)

# comparison
AICc(lvul_fixed, lvul_fixed1, lvul_fixed2, lvul_fixed3, lvul_fixed4, lvul_fixed5,
     lvul_mixed, lvul_mixed1, lvul_mixed2, lvul_mixed3, lvul_mixed4, lvul_mixed5)
```

```{r final model}
lvul_fixed <-  lm(data = df_lvul_performance[df_lvul_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + height23)

# normality
qqnorm(residuals(lvul_fixed))
qqline(residuals(lvul_fixed))

shapiro.test(residuals(lvul_fixed))

# coefficients
summary(lvul_fixed)
Anova(lvul_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(lvul_fixed)

# post-hoc
lvul_emmeans <- emmeans(lvul_fixed, ~ treatment + habitat,
                        type = "response")
summary(lvul_emmeans)
lvul_emmeans3 <- cld(lvul_emmeans, Letters = letters) # significance letters

lvul_pairs <- pairs(lvul_emmeans, adjust = "tukey")
```

#### Figure

```{r}
# bind emmeans of performance measurements to summary
emm_df <- full_join(lvul_perf_sum[lvul_perf_sum$performance == "biomass", ],
                    lvul_emmeans3,
                    by = c("habitat", "treatment"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)


fig_lvulmass <- lvul_perf_sum %>% 
  filter(performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~performance,
             labeller = as_labeller(c("biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL, 
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.1))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none")

# ggsave(here::here("4.4_Soil_Depth/plots/fig_lvultotalmass.png"), fig_lvulmass,
       # width = 12, height = 7)
```

### Combined performance figure

```{r}
# add performance labels
lvul_emmeans1$performance <- "height"
lvul_emmeans2$performance <- "ramets"
lvul_emmeans3$performance <- "biomass"

# bind emmeans of performance measurements
lvul_emmeans <- rbind(lvul_emmeans1, lvul_emmeans2, lvul_emmeans3)

# bind to summary
emm_df <- full_join(lvul_perf_sum,
                    lvul_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0, releve factors
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group) %>% 
  mutate(performance = factor(performance, levels = c("height", "ramets", "biomass")))


# figure
fig_lvulperfmass <- lvul_perf_sum %>% 
  mutate(performance = factor(performance, levels = c("height", "ramets", "biomass"))) %>% 
  filter(performance == "height" | performance == "ramets" | performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~ performance,
             labeller = as_labeller(c("height" = "Mean height (cm)",
                                      "ramets" = "Mean number of ramets",
                                      "biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL, 
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.1))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside")
```


## *P. major*

### Survival

```{r total survival}
# figure
fig_pmajtotalsurvival <- survival_sum %>% 
  group_by(habitat, treatment) %>% 
  filter(species == "Pmaj") %>% 
  
  # plot
  ggplot(aes(x = treatment,
             y = propsurvival,
             fill = habitat)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste("n =", alive)),
            position = position_dodge(0.9),
            vjust = -0.5) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Total number of surviving transplants") +
  scale_fill_brewer(name = "Soil Origin",
                    palette = "Accent",
                    labels = c("Dump", "Tundra")) +
  theme_vz()

# ggsave(here::here("4.4_Soil_Depth/plots/fig_pmajtotalsurvival.png"), fig_pmajtotalsurvival,
#        width = 12, height = 7)
```

#### GLM
```{r}
# glm
survival_fixed <- glm(survival ~ treatment + habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Pmaj", ],
                      subset = year_survivalf != "23")

survival_fixed1 <- glm(survival ~ treatment * habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Pmaj", ],
                      subset = year_survivalf != "23")

summary(survival_fixed)
Anova(survival_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")

survival_mixed <- glmer(survival ~ treatment + habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Pmaj", ],
                        subset = year_survivalf != "23")

survival_mixed1 <- glmer(survival ~ treatment * habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Pmaj", ],
                        subset = year_survivalf != "23")

survival_mixed2 <- glmer(survival ~ habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Pmaj", ],
                        subset = year_survivalf != "23")

survival_mixed3 <- glmer(survival ~ treatment + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Pmaj", ],
                        subset = year_survivalf != "23")

# model comparison
AICc(survival_fixed, survival_fixed1, survival_mixed, survival_mixed1, survival_mixed2, survival_mixed3)
```

```{r}
survival_mixed <- glm(survival ~ treatment + habitat,
                        family = "binomial",
                        data = df_survival[df_survival$species == "Pmaj", ],
                        subset = year_survivalf != "23")

summary(survival_mixed)
Anova(survival_mixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")


# post-hoc
surv_pairs <- emmeans(survival_fixed, ~ treatment + habitat,
                      type = "response")
summary(surv_pairs)
cld(surv_pairs, Letters = letters)

pairs(surv_pairs, adjust = "tukey", reverse = "TRUE")
```

### Performance

```{r normality}
# normality of height
shapiro.test(df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ]$values)
hist(df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ]$values)

# normality of ramets
shapiro.test(df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ]$values)
hist(df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ]$values)

# normality of biomass
shapiro.test(df_pmaj_performance[df_pmaj_performance$performance == "biomass", ]$values)
hist(df_pmaj_performance[df_pmaj_performance$performance == "biomass", ]$values)
```

```{r transformation}
df_pmaj_performance$values_transformed <- log(df_pmaj_performance$values)
```

```{r Pmaj performance summary}
pmaj_perf_sum <- df_pmaj_performance %>% 
  group_by(habitat, treatment, species, performance) %>% 
  filter(!is.na(values) &
           (performance == "leaflength" | performance == "numleaves" | performance == "biomass")) %>% 
  summarize(mean = mean(values),
            n = n(),
            sd = sd(values),
            se = sd/sqrt(n)) %>%
  mutate(mean = replace_na(mean, 0),
         se = replace_na(se, 0))
```


#### LMM

Need to use transformed values for numleaves.

```{r leaflength models}
# lm
pmaj_fixed1 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                 values ~ treatment * habitat + leaflength23)

pmaj_fixed2 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                 values ~ treatment + habitat + leaflength23)

pmaj_fixed3 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                 values ~ treatment * habitat)

pmaj_fixed <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                 values ~ treatment + habitat)

# lmer
pmaj_mixed1 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                   values ~ treatment * habitat + leaflength23 + (1|site),
                   REML = F)

pmaj_mixed2 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                   values ~ treatment + habitat + leaflength23 + (1|site),
                   REML = F)

pmaj_mixed3 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                   values ~ treatment * habitat + (1|site),
                   REML = F)

pmaj_mixed <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                   values ~ treatment + habitat + (1|site),
                   REML = F)

# comparison
AICc(pmaj_fixed, pmaj_fixed1, pmaj_fixed2, pmaj_fixed3,
     pmaj_mixed, pmaj_mixed1, pmaj_mixed2, pmaj_mixed3)
```


```{r final model}
pmaj_fixed <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "leaflength", ],
                 values ~ treatment + habitat)

# check normality
qqnorm(residuals(pmaj_fixed))
qqline(residuals(pmaj_fixed))

shapiro.test(residuals(pmaj_fixed))

# coefficients
summary(pmaj_fixed)
Anova(pmaj_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(pmaj_fixed)

# post-hoc
pmaj_emmeans <- emmeans(pmaj_fixed, ~ treatment * habitat,
                        type = "response")
summary(pmaj_emmeans)
pmaj_emmeans1 <- cld(pmaj_emmeans, Letters = letters) # significance letters

pairs(pmaj_emmeans, adjust = "tukey")
```

```{r numleaves models}
# lm
pmaj_fixed1 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                 values_transformed ~ treatment * habitat + leaflength23)

pmaj_fixed2 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat + leaflength23)

pmaj_fixed3 <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                 values_transformed ~ treatment * habitat)

pmaj_fixed <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat)


# lmer
pmaj_mixed1 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                   values_transformed ~ treatment * habitat + leaflength23 + (1|site),
                   REML = F)

pmaj_mixed2 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                   values_transformed ~ treatment + habitat + leaflength23 + (1|site),
                   REML = F)

pmaj_mixed3 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                   values_transformed ~ treatment * habitat + (1|site),
                   REML = F)

pmaj_mixed <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                   values_transformed ~ treatment + habitat + (1|site),
                   REML = F)

# comparison
AICc(pmaj_fixed, pmaj_fixed1, pmaj_fixed2, pmaj_fixed3,
     pmaj_mixed, pmaj_mixed1, pmaj_mixed2, pmaj_mixed3)
```


```{r final model}
pmaj_fixed <- lm(data = df_pmaj_performance[df_pmaj_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat)

# check normality
qqnorm(residuals(pmaj_fixed))
qqline(residuals(pmaj_fixed))

shapiro.test(residuals(pmaj_fixed))

# coefficients
summary(pmaj_fixed)
Anova(pmaj_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(pmaj_fixed)

# post-hoc
pmaj_emmeans <- emmeans(pmaj_fixed, ~ treatment * habitat,
                        type = "response")
summary(pmaj_emmeans)
pmaj_emmeans2 <- cld(pmaj_emmeans, Letters = letters) # significance letters

pairs(pmaj_emmeans, adjust = "tukey")
```

#### Figure
```{r Pmaj performance plot}
# add performance labels
pmaj_emmeans1$performance <- "leaflength"
pmaj_emmeans2$performance <- "numleaves"

# bind emmeans of performance measurements
pmaj_emmeans <- rbind(pmaj_emmeans1, pmaj_emmeans2)

# bind to summary
emm_df <- full_join(pmaj_perf_sum[pmaj_perf_sum$performance != "biomass", ],
                    pmaj_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)

fig_pmajperf <- pmaj_perf_sum %>% 
  filter(performance == "leaflength" | performance == "numleaves") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  geom_col(aes(fill = habitat), position = position_dodge2(preserve = "single"),
           colour = "black") +
  facet_wrap(~ performance,
             labeller = as_labeller(c("leaflength" = "Mean leaf length (cm)",
                                      "numleaves" = "Mean number of leaves")),
             scales = "free_y",
             strip.position = "left") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_x_discrete(name = "", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL,
                     limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  # theme_darkclassic() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none"
        )

# ggsave(here::here("4.4_Soil_Depth/plots/fig_pmajperf.png"), fig_pmajperf,
#        width = 12, height = 7)
```


### Flower

```{r flower models}
# lm
pmaj_fixed1 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                  values ~ treatment * habitat + leaflength23)

pmaj_fixed2 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                  values ~ treatment + habitat + leaflength23)

pmaj_fixed3 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                  values ~ treatment * habitat)

pmaj_fixed <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                  values ~ treatment + habitat)

# lmer
pmaj_mixed1 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                    values ~ treatment * habitat + leaflength23 + (1|site),
                    REML = F)

pmaj_mixed <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                    values ~ treatment + habitat + leaflength23 + (1|site),
                    REML = F)

pmaj_mixed2 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                    values ~ treatment * habitat + (1 | site),
                    REML = F)

pmaj_mixed3 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                    values ~ treatment + habitat + (1 | site), 
                    REML = F)

# comparison
AICc(pmaj_fixed, pmaj_fixed1, pmaj_fixed2, pmaj_fixed3,
     pmaj_mixed, pmaj_mixed1, pmaj_mixed2, pmaj_mixed3)
```

```{r final model}
pmaj_fixed <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "flower", ],
                  values ~ treatment + habitat)

summary(pmaj_fixed)
Anova(pmaj_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(pmaj_mixed)

# post-hoc
pmaj_emmeans <- emmeans(pmaj_mixed, ~ treatment + habitat,
                        type = "response")
summary(pmaj_emmeans)
pmaj_emmeans3 <- cld(pmaj_emmeans, Letters = letters) # significance letters

pairs(pmaj_emmeans, adjust = "tukey")
```

### Biomass

```{r biomass models}
# lm
pmaj_fixed1 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + leaflength23)

pmaj_fixed2 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23)

pmaj_fixed3 <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat)

pmaj_fixed <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat)


# lmer
pmaj_mixed1 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                    values_transformed ~ treatment * habitat + leaflength23 + (1|site),
                    REML = F)

pmaj_mixed <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                    values_transformed ~ treatment + habitat + leaflength23 + (1|site),
                    REML = F)

pmaj_mixed2 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                    values_transformed ~ treatment * habitat + (1 | site),
                    REML = F)

pmaj_mixed3 <- lmer(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                    values_transformed ~ treatment + habitat + (1 | site), 
                    REML = F)

# comparison
AICc(pmaj_fixed, pmaj_fixed1, pmaj_fixed2, pmaj_fixed3,
     pmaj_mixed, pmaj_mixed1, pmaj_mixed2, pmaj_mixed3)
```

```{r final model}
pmaj_fixed <-  lm(data = df_pmaj_performance[df_pmaj_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat)

# normality
qqnorm(residuals(pmaj_fixed))
qqline(residuals(pmaj_fixed))

shapiro.test(residuals(pmaj_fixed))

# coefficients
summary(pmaj_fixed)
Anova(pmaj_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(pmaj_fixed)


# post-hoc
pmaj_emmeans <- emmeans(pmaj_fixed, ~ treatment + habitat,
                        type = "response")
summary(pmaj_emmeans)
pmaj_emmeans3 <- cld(pmaj_emmeans, Letters = letters) # significance letters

pairs(pmaj_emmeans, adjust = "tukey")
```

#### Figure

```{r}
# bind emmeans of performance measurements to summary
emm_df <- full_join(pmaj_perf_sum[pmaj_perf_sum$performance == "biomass", ],
                    pmaj_emmeans3,
                    by = c("habitat", "treatment"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)

fig_pmajmass <- pmaj_perf_sum %>% 
  filter(performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~performance,
             labeller = as_labeller(c("biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat), position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL, 
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.1))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none")

# ggsave(here::here("4.4_Soil_Depth/plots/fig_pmajtotalmass.png"), fig_pmajmass,
#        width = 12, height = 7)
```

### Combined performance figure

```{r}
# add performance labels
pmaj_emmeans1$performance <- "leaflength"
pmaj_emmeans2$performance <- "numleaves"
pmaj_emmeans3$performance <- "biomass"

# bind emmeans of performance measurements
pmaj_emmeans <- rbind(pmaj_emmeans1, pmaj_emmeans2, pmaj_emmeans3)

# bind to summary
emm_df <- full_join(pmaj_perf_sum,
                    pmaj_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0, releve factors
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group) %>% 
  mutate(performance = factor(performance, levels = c("leaflength", "numleaves", "biomass")))


# figure
fig_pmajperfmass <- pmaj_perf_sum %>% 
  mutate(performance = factor(performance, levels = c("leaflength", "numleaves", "biomass"))) %>% 
  filter(performance == "leaflength" | performance == "numleaves" | performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~ performance,
             labeller = as_labeller(c("leaflength" = "Mean leaf length (cm)",
                                      "numleaves" = "Mean number of leaves",
                                      "biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL, 
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.1))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside")
```


## *T. officinale*

### Survival

```{r total survival}
# figure
fig_tofftotalsurvival <- survival_sum %>% 
  group_by(habitat, treatment) %>% 
  filter(species == "Toff") %>% 
  
  # plot
  ggplot(aes(x = treatment,
             y = propsurvival,
             fill = habitat)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste("n =", alive)),
            position = position_dodge(0.9),
            vjust = -0.5) +
  scale_x_discrete(name = "Soil depth (cm)",
                   labels = c("Control", "5", "20")) +
  scale_y_continuous(name = "Total number of surviving transplants") +
  scale_fill_brewer(name = "Soil Origin",
                    palette = "Accent",
                    labels = c("Dump", "Tundra")) +
  theme_vz()

# ggsave(here::here("4.4_Soil_Depth/plots/fig_tofftotalsurvival.png"), fig_tofftotalsurvival,
#        width = 12, height = 7)
```

#### GLM

```{r}
# glm
survival_fixed <- glm(survival ~ treatment + habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Toff", ],
                      subset = year_survivalf != "23")

survival_fixed1 <- glm(survival ~ treatment * habitat,
                      family = "binomial",
                      data = df_survival[df_survival$species == "Toff", ],
                      subset = year_survivalf != "23")

summary(survival_fixed)
Anova(survival_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")

survival_mixed <- glmer(survival ~ treatment + habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Toff", ],
                        subset = year_survivalf != "23")

survival_mixed1 <- glmer(survival ~ treatment * habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Toff", ],
                        subset = year_survivalf != "23")

survival_mixed2 <- glmer(survival ~ habitat + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Toff", ],
                        subset = year_survivalf != "23")

survival_mixed3 <- glmer(survival ~ treatment + (1|site),
                        family = "binomial",
                        data = df_survival[df_survival$species == "Toff", ],
                        subset = year_survivalf != "23")

# model comparison
AICc(survival_fixed, survival_fixed1, survival_mixed, survival_mixed1, survival_mixed2, survival_mixed3)
```

```{r}
survival_mixed <- glm(survival ~ treatment + habitat,
                        family = "binomial",
                        data = df_survival[df_survival$species == "Toff", ],
                        subset = year_survivalf != "23")

summary(survival_mixed)
Anova(survival_mixed, p.adjust.method = TRUE, type = "3", test.statistic = "Wald")


# post-hoc
surv_pairs <- emmeans(survival_fixed, ~ treatment + habitat,
                      type = "response")
summary(surv_pairs)
cld(surv_pairs, Letters = letters)

pairs(surv_pairs, adjust = "tukey", reverse = "TRUE")
```

### Performance

```{r normality}
# normality of height
shapiro.test(df_toff_performance[df_toff_performance$performance == "leaflength", ]$values)
hist(df_toff_performance[df_toff_performance$performance == "leaflength", ]$values)

# normality of ramets
shapiro.test(df_toff_performance[df_toff_performance$performance == "numleaves", ]$values)
hist(df_toff_performance[df_toff_performance$performance == "numleaves", ]$values)

# normality of biomass
shapiro.test(df_toff_performance[df_toff_performance$performance == "biomass", ]$values)
hist(df_toff_performance[df_toff_performance$performance == "biomass", ]$values)
```

```{r transformation}
df_toff_performance$values_transformed <- log(df_toff_performance$values)
```

```{r Toff performance summary}
toff_perf_sum <- df_toff_performance %>% 
  group_by(habitat, treatment, species, performance) %>% 
  filter(!is.na(values) &
           (performance == "leaflength" | performance == "numleaves" | performance == "biomass")) %>% 
  summarize(mean = mean(values),
            n = n(),
            sd = sd(values),
            se = sd/sqrt(n)) %>%
  mutate(mean = replace_na(mean, 0),
         se = replace_na(se, 0))
```

#### LMM

Need to use transformed values for leaflength and numleaves.

```{r leaflength models}
toff_fixed1 <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment * habitat + leaflength23 + taproot23)

toff_fixed2 <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment + habitat + leaflength23 + taproot23)

toff_fixed3 <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment * habitat + leaflength23)

toff_fixed <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment + habitat + leaflength23)

toff_fixed4 <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment * habitat + taproot23)

toff_fixed5 <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment + habitat + taproot23)

# lmer
toff_mixed1 <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment * habitat + leaflength23 + taproot23 + (1|site),
                   REML = F)

toff_mixed2 <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment + habitat + leaflength23 + taproot23 + (1|site),
                   REML = F)

toff_mixed3 <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment * habitat + leaflength23 + (1|site),
                   REML = F)

toff_mixed <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment + habitat + leaflength23 + (1|site),
                   REML = F)

toff_mixed4 <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment * habitat + taproot23 + (1|site),
                   REML = F)

toff_mixed5 <- lmer(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                   values_transformed ~ treatment + habitat + taproot23 + (1|site),
                   REML = F)

# comparison
AICc(toff_fixed, toff_fixed1, toff_fixed2, toff_fixed3, toff_fixed4, toff_fixed5,
     toff_mixed, toff_mixed1, toff_mixed2, toff_mixed3, toff_mixed4, toff_mixed5)
```

```{r final model}
toff_fixed <- lm(data = df_toff_performance[df_toff_performance$performance == "leaflength", ],
                 values_transformed ~ treatment + habitat + leaflength23)

# check normality
qqnorm(residuals(toff_fixed))
qqline(residuals(toff_fixed))

shapiro.test(residuals(toff_fixed))

# coefficients
summary(toff_fixed)
Anova(toff_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(toff_fixed)

# post-hoc
toff_emmeans <- emmeans(toff_fixed, ~ treatment + habitat,
                        type = "response")
summary(toff_emmeans)
toff_emmeans1 <- cld(toff_emmeans, Letters = letters) # significance letters

pairs(toff_emmeans, adjust = "tukey")
```

```{r numleaves models}
toff_fixed1 <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment * habitat + leaflength23 + taproot23)

toff_fixed2 <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat + leaflength23 + taproot23)

toff_fixed3 <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment * habitat + leaflength23)

toff_fixed <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat + leaflength23)

toff_fixed4 <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment * habitat + taproot23)

toff_fixed5 <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat + taproot23)

# lmer
toff_mixed1 <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment * habitat + leaflength23 + taproot23 + (1|site),
                   REML = F)

toff_mixed2 <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment + habitat + leaflength23 + taproot23 + (1|site),
                   REML = F)

toff_mixed3 <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment * habitat + leaflength23 + (1|site),
                   REML = F)

toff_mixed <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment + habitat + leaflength23 + (1|site),
                   REML = F)

toff_mixed4 <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment * habitat + taproot23 + (1|site),
                   REML = F)

toff_mixed5 <- lmer(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                   values_transformed ~ treatment + habitat + taproot23 + (1|site),
                   REML = F)

# comparison
AICc(toff_fixed, toff_fixed1, toff_fixed2, toff_fixed3, toff_fixed4, toff_fixed5,
     toff_mixed, toff_mixed1, toff_mixed2, toff_mixed3, toff_mixed4, toff_mixed5)
```


```{r final model}
toff_fixed <- lm(data = df_toff_performance[df_toff_performance$performance == "numleaves", ],
                 values_transformed ~ treatment + habitat + leaflength23)

# check normality
qqnorm(residuals(toff_fixed))
qqline(residuals(toff_fixed))

shapiro.test(residuals(toff_fixed))

# coefficients
summary(toff_fixed)
Anova(toff_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(toff_fixed)

# post-hoc
toff_emmeans <- emmeans(toff_fixed, ~ treatment + habitat,
                        type = "response")
summary(toff_emmeans)
toff_emmeans2 <- cld(toff_emmeans, Letters = letters) # significance letters

pairs(toff_emmeans, adjust = "tukey")
```

#### Figure

```{r Toff performance plot}
# add performance labels
toff_emmeans1$performance <- "leaflength"
toff_emmeans2$performance <- "numleaves"

# bind emmeans of performance measurements
toff_emmeans <- rbind(toff_emmeans1, toff_emmeans2)

# bind to summary
emm_df <- full_join(toff_perf_sum[toff_perf_sum$performance != "biomass", ],
                    toff_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)

fig_toffperf <- toff_perf_sum %>% 
  filter(performance == "leaflength" | performance == "numleaves") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  geom_col(aes(fill = habitat), position = position_dodge2(preserve = "single"),
           colour = "black") +
  facet_wrap(~ performance,
             labeller = as_labeller(c("leaflength" = "Mean leaf length (cm)",
                                      "numleaves" = "Mean number of leaves")),
             scales = "free_y",
             strip.position = "left") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL,
                     limits = c(0, NA), expand = expansion(mult = c(0, 0.2))) + 
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  # theme_darkclassic() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none"
        )

# ggsave(here::here("4.4_Soil_Depth/plots/fig_toffperf.png"), fig_toffperf,
#        width = 12, height = 7)
```

### Flower
```{r flower models}
toff_fixed1 <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + leaflength23 + taproot23)

toff_fixed2 <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + leaflength23 + taproot23)

toff_fixed3 <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + leaflength23)

toff_fixed <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + leaflength23)

toff_fixed4 <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + taproot23)

toff_fixed5 <-  lm(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + taproot23)


# lmer
toff_mixed1 <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + leaflength23 + taproot23 + (1|site),
                  REML = F)

toff_mixed2 <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + leaflength23 + taproot23 + (1|site),
                  REML = F)

toff_mixed3 <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + leaflength23 + (1|site),
                  REML = F)

toff_mixed <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + leaflength23 + (1|site),
                  REML = F)

toff_mixed4 <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment * habitat + taproot23 + (1|site),
                  REML = F)

toff_mixed5 <- lmer(data = df_toff_performance[df_toff_performance$performance == "flower", ],
                  values ~ treatment + habitat + taproot23 + (1|site),
                  REML = F)

# comparison
AICc(toff_fixed, toff_fixed1, toff_fixed2, toff_fixed3, toff_fixed4, toff_fixed5,
     toff_mixed, toff_mixed1, toff_mixed2, toff_mixed3, toff_mixed4, toff_mixed5)
```


```{r final model}
summary(toff_fixed)
Anova(toff_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(toff_mixed)

# post-hoc
toff_emmeans <- emmeans(toff_mixed, ~ treatment + habitat,
                        type = "response")
summary(toff_emmeans)
toff_emmeans3 <- cld(toff_emmeans, Letters = letters) # significance letters

pairs(toff_emmeans, adjust = "tukey")
```


### Biomass

```{r biomass models}
toff_fixed1 <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + leaflength23 + taproot23)

toff_fixed2 <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23 + taproot23)

toff_fixed3 <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + leaflength23)

toff_fixed <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23)

toff_fixed4 <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + taproot23)

toff_fixed5 <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + taproot23)

# lmer
toff_mixed1 <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + leaflength23 + taproot23 + (1|site),
                  REML = F)

toff_mixed2 <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23 + taproot23 + (1|site),
                  REML = F)

toff_mixed3 <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + leaflength23 + (1|site),
                  REML = F)

toff_mixed <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23 + (1|site),
                  REML = F)

toff_mixed4 <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment * habitat + taproot23 + (1|site),
                  REML = F)

toff_mixed5 <- lmer(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + taproot23 + (1|site),
                  REML = F)

# comparison
AICc(toff_fixed, toff_fixed1, toff_fixed2, toff_fixed3, toff_fixed4, toff_fixed5,
     toff_mixed, toff_mixed1, toff_mixed2, toff_mixed3, toff_mixed4, toff_mixed5)
```


```{r final model}
toff_fixed <-  lm(data = df_toff_performance[df_toff_performance$performance == "biomass", ],
                  values_transformed ~ treatment + habitat + leaflength23)

# normality
qqnorm(residuals(toff_fixed))
qqline(residuals(toff_fixed))

shapiro.test(residuals(toff_fixed))

# coefficients
summary(toff_fixed)
Anova(toff_fixed, p.adjust.method = TRUE, type = "3", test.statistic = "F")

# r2
r.squaredGLMM(toff_fixed)

# post-hoc
toff_emmeans <- emmeans(toff_fixed, ~ treatment + habitat,
                        type = "response")
summary(toff_emmeans)
toff_emmeans3 <- cld(toff_emmeans, Letters = letters) # significance letters

pairs(toff_emmeans, adjust = "tukey")
```

#### Figure

```{r}
# bind emmeans of performance measurements to summary
emm_df <- full_join(toff_perf_sum[toff_perf_sum$performance == "biomass", ],
                    toff_emmeans3,
                    by = c("habitat", "treatment"))

# remove measurements that are 0
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group)


fig_toffmass <- toff_perf_sum %>% 
  filter(performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~performance,
             labeller = as_labeller(c("biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat), position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL,
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.2))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none")

# ggsave(here::here("4.4_Soil_Depth/plots/fig_tofftotalmass.png"), fig_toffmass,
#        width = 12, height = 7)
```

### Combined performance figure

```{r}
# add performance labels
toff_emmeans1$performance <- "leaflength"
toff_emmeans2$performance <- "numleaves"
toff_emmeans3$performance <- "biomass"

# bind emmeans of performance measurements
toff_emmeans <- rbind(toff_emmeans1, toff_emmeans2, toff_emmeans3)

# bind to summary
emm_df <- full_join(toff_perf_sum,
                    toff_emmeans,
                    by = c("habitat", "treatment", "performance"))

# remove measurements that are 0, releve factors
emm_df <- emm_df %>% 
  mutate(group = ifelse(mean == 0, "", .group)) %>% 
  dplyr::select(-.group) %>% 
  mutate(performance = factor(performance, levels = c("leaflength", "numleaves", "biomass")))


# figure
fig_toffperfmass <- toff_perf_sum %>% 
  mutate(performance = factor(performance, levels = c("leaflength", "numleaves", "biomass"))) %>% 
  filter(performance == "leaflength" | performance == "numleaves" | performance == "biomass") %>% 
  ggplot(aes(x = treatment, y = mean)) +
  facet_wrap(~ performance,
             labeller = as_labeller(c("leaflength" = "Mean leaf length (cm)",
                                      "numleaves" = "Mean number of leaves",
                                      "biomass" = "Mean biomass (mg)")),
             scales = "free_y",
             strip.position = "left") +
  geom_col(aes(fill = habitat),
           position = position_dodge2(preserve = "single"),
           colour = "black") +
  geom_errorbar(aes(x = treatment, ymin = mean - se, ymax = mean + se),
                position = position_dodge2(padding = 0.5),
                colour = "black") +
  scale_fill_brewer(name = "Habitat",
                    palette = "Accent",
                    labels = c("Boreal", "Tundra")) +
  scale_x_discrete(name = "Soil depth (cm)", labels = c("Control", "5", "20")) +
  scale_y_continuous(name = NULL, 
                     limits = c(-0.01, NA), expand = expansion(mult = c(0, 0.1))) + 
  geom_text(data = emm_df,
            aes(y = mean + se, label = group,
                vjust = -1),
            position = position_dodge2(width = 0.9)) +
  theme_vz() +
  theme(strip.background = element_blank(),
        strip.placement = "outside")
```


# Combined Figure

Survival & seed
```{r}
fig_survseed <- fig_sppsurvival + fig_seedbank +
  plot_layout(guides = "keep", nrow = 2) +
  plot_annotation(tag_levels = c("A", "B")) & 
  theme()
```


Performance
```{r}
fig_sppperfmass <- fig_lvulperfmass +
  fig_pmajperfmass +
  fig_toffperfmass +
  plot_layout(guides = "collect", nrow = 3) +
  plot_annotation(tag_levels = c("A", "B", "C")) 
```




